---
title: "Lab6_ESM244_Mireia"
author: "Mireia Valle"
date: "13/2/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warnings = FALSE)
```

## Loading libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(raster)
library(sf)
library(tmap)
library(tmaptools)
#install.packages("gstat")
library(gstat)
```

## Gran Canyon GeoTIFF

```{r}
gc_dem <- raster(here("data", "gc_dem.tif"))

# Look at it using plot
plot(gc_dem)

# Check the CRS
gc_dem@crs


# Check the extent (bounding box)

gc_dem@extent

# Creating a wgs84 w/latlong

wgs84 = "+proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs"

# Reproject
gc_reproj = projectRaster(gc_dem, crs = wgs84, method = "bilinear")

# Then check: aha, now degrees we're used to
gc_reproj@extent
```

#### Crop the raster to a smaller area: 
```{r}
bounds <- as(extent(-112.4, -112.0, 36.1, 36.3), 'SpatialPolygons')

# Make the projection for "bounds" the same as for "gc_reproj":
crs(bounds) <- crs(gc_reproj)

# Then crop gc_reproj by the new bounds polygon: 
gc_crop <- crop(gc_reproj, bounds)

# Look at it:
plot(gc_crop)
```

Want to resample? Use `raster::aggregate()` to create lower res (larger cell) rasters.

See ?aggregrate (default is mean, `fact` is number of cells in each direction; can set 2 if wanting x/y to differ for aggregation)
```{r}
# Aggregate: 
gc_agg <- aggregate(gc_crop, fact = 10)

# Then look at it: 
plot(gc_agg)
```

### Get safe and sound back in ggplot:
```{r}
# First, convert to a data frame:
gc_df <- as.data.frame(gc_agg, xy = TRUE)

# That `xy = TRUE` is important: retains the lat/lon information! 

# View(gc_df)

ggplot(data = gc_df, aes(x = x, y = y)) +
  geom_raster(aes(fill = layer)) +
  coord_quickmap() +
  theme_minimal() +
  scale_fill_gradientn(colors = c("purple",
                                  "magenta",
                                  "orange",
                                  "yellow", 
                                  "white")
                       )
```

Let's say we know that in this region, a certain species will only grow between 1000 and 1500 ft elevation. Create a subset of gc_crop that includes that habitat:
```{r}
# First, make a copy
gc_hab <- gc_crop

# Set any cells outside of 1000 - 1500 to NA
gc_hab[gc_hab > 1500 | gc_hab < 1000] <- NA

# Plot.
plot(gc_hab)

# Cool! 
```
Let's make a bit nicer map with `tmap`: 
```{r}
tmap_mode("view") # Set to interactive viewing

# Make tmap: 
tm_shape(gc_hab) +
  tm_raster(legend.show = FALSE, palette = "plasma")

# In console, run tmaptools::palette_explorer() to view a Shiny app with other palettes!
```

## Kriging Kansas Rain

#### Read in the KS counties shapefile with `read_sf`:
```{r}
ks_counties <- read_sf(here("data",
                            "ks_counties",
                            "ks_counties_shapefile.shp"))

# View(ks_counties)
```

#### Do some exploring:
```{r}
# Base plot
plot(ks_counties)

# Check CRS:
st_crs(ks_counties) # hmmmm none...guess we should set one!

# Set to EPSG 4326 (WGS84 datum):
st_crs(ks_counties) <- 4326

# Now check again:
st_crs(ks_counties)

# And replot:
plot(ks_counties)

# That looks more like Kansas. 
```

#### Now, read in the KS rainfall data (this was from a single rainy day in 2017): 
```{r}
ks_rain <- read_csv(here("data","ks_rain.csv")) %>% 
  clean_names()
```

But currently, R has no idea that these are spatial points. We'll convert it using `sf::st_as_sf()`:

```{r}
ks_sf  <-  st_as_sf(ks_rain, coords = c("lon", "lat"), 
                 crs = 4326)

# View(ks_sf) (rainfall = amt)
plot(ks_sf)

# Or in ggplot:
ggplot() +
  geom_sf(data = ks_counties) +
  geom_sf(data = ks_sf, aes(color = amt,
              size = amt),
          show.legend = FALSE) +
  coord_sf() +
  scale_color_gradient(low = "yellow", high = "red") +
  theme_void()
```

## Kriging to predict rainfall:

```{r}
#?as_Spatial
ks_sp <- as_Spatial(ks_sf)
class(ks_sp)
```

### Make a spatial pixels grid that we'll make predictions over: 

```{r}
bbox(ks_sp) #to check bounding box of the spatial points
lat <- seq(37, 40, length.out = 200)
long <- seq(-94.6, -102, length.out = 200)

# Now we're going to make this into a spatial grid: 

grid <- expand.grid(lon = long, lat = lat)

grid_sf <- st_as_sf(grid, coords = c("lon", "lat"), crs = 4326)
grid_sp <- as_Spatial(grid_sf)
```

### Make a variogram

```{r}
ks_vgm <- variogram(amt ~ 1, data = ks_sp)
plot(ks_vgm)

#My estimates for variogram parameters
#nugget = 0.2
#still = 0.8
#range = 200

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2, psill = 0.8, range = 200, model = "Sph"))

#other models = "Gau", "Exp"

plot(ks_vgm, ks_vgm_fit)

ks_vgm_fit

```
Now, kriging! 
```{r}
ks_krige <- krige(amt ~ 1, ks_sp, grid_sp, model=ks_vgm_fit)
ks_krige


spplot(ks_krige, "var1.pred")

spplot(ks_krige, "var1.var")
```

### Make a data frame of the kridged predictions

```{r}
ks_df <- data.frame(ks_krige@data["var1.pred"],
                    ks_krige@data["var1.var"],
                    ks_krige@coords) %>% 
  rename(longitude = coords.x1,
         latitude = coords.x2)

#Converting this df to a sf object

rain_sf <- st_as_sf(ks_df, coords = c("longitude","latitude"))
st_crs(rain_sf) <- 4326
```

```{r}

# Get Kansas outline to crop: 
ks <- read_sf(dsn = here("data","states"), 
              layer = "cb_2017_us_state_20m") %>% 
  dplyr::select(NAME) %>% 
  filter(NAME == "Kansas") %>% 
  st_transform(crs = 4326)

plot(ks)

```

```{r}
# Find the intersection of the two: 
rain_sf_ks <- st_intersection(rain_sf, ks)

ggplot(rain_sf_ks) +
  geom_sf(aes(color = var1.pred)) +
  scale_color_gradientn(colors = c("white","yellow","magenta","purple")) +
  theme_minimal()

ggplot(rain_sf_ks) +
  geom_sf(aes(color = var1.var)) +
  geom_sf(data = ks_sf, aes(color = amt,
              size = amt)) +
  theme_minimal()

```
